Division and Slope Factorization of p-Adic Polynomials

### 2016-02-01 polynomials/muldivfact.tex
Small improvements

\category{I.1.2}{Computing Methodologies}{Symbolic and Algebraic Manipulation -- \emph{Algebraic Algorithms}}
\terms{Algorithms, Theory}
\keywords{}

@@

Polynomial factorization over local fields has seen a great deal of progress recently
\citelist{\cite{pauli:10a}, \cite{guardia-nart-pauli:12a}, \cite{guardia-montes-nart:08a}, \cite{montes:99a}}
following an algorithm of Montes.  Slope factorization provides a subroutine
in such algorithms \cite[Section~2]{pauli:10a}.  For the most difficult inputs, it is not the dominant
contributor to the runtime of the algorithm, but in some circumstances it will be.
We underline moreover that the methods introduced in this paper extend 
partially to the noncommutative setting and appear this way as an
essential building block in several decomposition algorithms of 
$p$-adic Galois representations and $p$-adic differential 
equations~\cite{caruso:16}.

@@

The aim of this section is to design efficient and stable algorithms 
for computing these factorizations: indeed the algorithm we obtain has 
a quasi-optimal complexity (compared to the size of the input 
polynomial) and outputs a result whose precision is (close to be) 
optimal. One important feature of our algorithm is its simplicity 
and its flexibility: in particular, unlike standard previous slope 
factorization algorithms, it does not require working with
fractional exponents.

@@

In the previous references, the factor $A$ defined above is always 
obtained \emph{via} a Newton iteration which takes place over a finite 
extension of $K$. We introduce here a variant of this Newton iteration 
whose main advantage is to be defined over $K$ itself.

>>>>

\category{I.1.2}{Computing Methodologies}{Symbolic and Algebraic Manipulation -- \emph{Algebraic Algorithms}}
\terms{Algorithms, Theory}
\keywords{Algorithms, $p$-adic precision, Newton polygon, factorization}

@@

Polynomial factorization over local fields has seen a great deal of progress recently
\citelist{\cite{pauli:10a} \cite{guardia-nart-pauli:12a} \cite{guardia-montes-nart:08a} \cite{montes:99a}}
following an algorithm of Montes.  Slope factorization provides a subroutine
in such algorithms \cite[Section~2]{pauli:10a}.  For the most difficult inputs, it is not the dominant
contributor to the runtime of the algorithm, but in some circumstances it will be.
We underline moreover that the methods introduced in this paper extend 
partially to the noncommutative setting and appear this way as an
essential building block in several decomposition algorithms of 
$p$-adic Galois representations and $p$-adic differential 
equations~\cite{caruso:16}.

@@

The aim of this section is to design efficient and stable algorithms 
for computing these factorizations. Precisely the algorithm we obtain 
has a quasi-optimal complexity (compared to the size of the input 
polynomial) and outputs a result whose precision is (close to be) 
optimal. Two of its important additional features are simplicity and 
flexibility.

@@

The factor $A$ defined above is usually obtained \emph{via} a Newton 
iteration after having prepared our polynomial by flattening the first 
slope using a change of variables involving possibly rational 
exponents. We introduce here a variant of this iteration which does not 
require the flattening step and is entirely defined over $K[X]$.
